---
title: "Performance"
output:
    html_document:
        toc: true
        toc_depth: 3
        toc_float: true
        number_sections: true
        theme: cerulean
        highlight: pygments
        code_folding: hide
        fig_width: 9
        fig_height: 3
        css: style.css
params:
    datadir: "./data"
    graphdir: "./graphs"
    macrodir: "./macros"
---

```{r echo = FALSE, warning=FALSE, message=FALSE}
## https://bookdown.org/yihui/rmarkdown-cookbook/source-script.html
source("analysis.R", local = knitr::knit_global())
```

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
url <- "https://rir-benchmarks.prl.fit.cvut.cz/job/"
```

# H3.1: Peformance of base Ř vs. Ř strict

This experiment shows the comparison of performance on our benchmarks
for the base version of Ř and Ř strict.

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
performanceExperiment <- list(
  `Ř` = "rsh-perf",
  `Ř strict` = "rsh-strict-perf")
```

# H3.2: Promise allocations

This experiment compares the number of promise allocations in Ř and Ř strict;
the baseline is GNUR.

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
promiseReductionExperiment <- list(
  `GNU R` = "gnur-promises",
  `Ř` = "rsh-promises",
  `Ř strict` = "rsh-strict-promises")
```

This experiment shows the impact on GC of base Ř and Ř strict, and the percentage
speedup due to GC.
We reuse the performanceExperiment data.

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
gcExperiment <- list(
  `Ř` = "rsh-gc",
  `Ř strict` = "rsh-strict-gc")
```

# H3.3: Optimizer impact comparing Ř strict and Ř strict w/o optimizer

Here we compare strict Ř performance running with and without the optimizer.
Without, all code is executed in the first tier bytecode interpreter.

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
optimizerExperiment <- list(
  `Ř` = "rsh-perf",
  `Ř no-opt` = "rsh-perf-noopt",
  `Ř strict` = "rsh-strict-perf",
  `Ř strict no-opt` = "rsh-strict-perf-noopt")
```

# Benchmark selection

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
selection <- c(
  "Bounce",
  "Mandelbrot",
  "Storage",
  "Flexclust",
  "Binarytrees",
  "Fasta",
  "Fastaredux",
  "Fannkuchredux",
  "Knucleotide",
  "Nbody",
  "Pidigits",
  "Regexdna",
  "Reversecomplement",
  "Spectralnorm"
)
```

# Helpers

```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
# fetch <- function(job, addWarm) {
#   dir.create("data", showWarnings=F)
#   setwd("data")
#   file <- paste0(job, ".csv")
#   if (!file.exists(file))
#     download.file(paste0(url, file), file)
#   res <- read.csv(file, header=TRUE, strip.white=TRUE)
#   if (addWarm)
#     res$warm <- c(rep(0L, COLD_RUNS), rep(1L, WARM_RUNS))
#   res$row <- 1:nrow(res)
#   setwd("..")
#   res
# }

fetch <- function(job, addWarm, warmup, iterations) {
  dir.create("data", showWarnings=F)
  setwd("data")
  file <- paste0(job, ".csv")
  if (!file.exists(file))
    stop("file not found")
  res <- read.csv(file, header=TRUE, strip.white=TRUE)
  if (addWarm)
    res$warm <- c(rep(0L, warmup), rep(1L, iterations - warmup))
  res$row <- 1:nrow(res)
  setwd("..")
  res
}

geometric.mean <- function(x, na.rm=TRUE) {
  if (is.null(nrow(x))) {
    exp(mean(log(x), na.rm=TRUE))
  } else {
    exp(apply(log(x), 2, mean, na.rm=na.rm))
  }
}

printResult <- function(name, result, digits=3, space=F) {
  name = gsub("μ", "mu", name)
  r = format(result, digits=digits)
  cat(paste0("\\newcommand{\\result",name,"}{",r))
  if (space)
    cat("\\xspace")
  cat("}\n")
}

remove_warmup <- function(data) {
  data[data$warm == 1,]
}

fetch_jobs <- function(jobs, addWarm = FALSE, warmup, iterations) {
  data <- NULL
  for (i in seq_along(jobs)) {
    job <- jobs[i]
    ex = names(job)
    d <- fetch(job, addWarm, warmup, iterations)
    d$experiment = ex
    if (is.null(data))
      data <- d
    else
      data <- merge(data, d, all=TRUE)
  }
  data
}

normalize <- function(data, removeNormalized, baseline) {
  data$speedup = 1
  for (b in unique(data$benchmark)) {
    if (missing(baseline))
      e1 = unique(data$experiment)[1]
    else
      e1 = baseline
    m = median(data[data$benchmark == b & data$experiment == e1 & data$warm == 1, ]$ms)
    data[data$benchmark == b, ]$speedup <-
      m/data[data$benchmark == b, ]$ms
    if (removeNormalized)
      data = data[-which(data$benchmark == b & data$experiment == e1), ]
  }
  data
}

shorten_suite <- function(s) {
  switch(s,
         "are-we-fast-r"="awf",
         "real_thing"="re",
         "real_thing_annotations"="reA",
         "shootout"="sht",
         "shootout_annotations"="shA",
         "simple"="μ")
}

printable_name <- function(s) {
  s <- gsub("_", " ", s)
  substr(s, 1, 1) <- toupper(substr(s, 1, 1))
  s
}

```



```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
H3.1 <- function() {
  perf <- fetch_jobs(performanceExperiment)
  perf$experiment = factor(perf$experiment, c("Ř", "Ř strict"))
  perf$benchmark_ = perf$benchmark
  perf$suite_short = lapply(perf$suite, shorten_suite)
  perf$benchmark = paste0(lapply(perf$benchmark, printable_name), "")
  sss <<- perf
  perf <- normalize(perf, T, "Ř")
  perf_ <- perf[perf$benchmark %in% selection & perf$warm == 1, ]

  perf_plot <-
    ggplot(perf_, aes(benchmark, speedup)) +
    geom_boxplot() +
    geom_hline(yintercept=1) +
    labs(y = "Speedup", x = "") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  save_graph(perf_plot, "rshStrictPerf")
}
H3.1()
```



```{r echo=FALSE, warning=FALSE, message=FALSE, eval=TRUE}
H3.2a <- function() {
  perf <- fetch_jobs(performanceExperiment)
  perf$experiment = factor(perf$experiment, c("Ř", "Ř-strict"))
  perf$benchmark_ = perf$benchmark
  perf$suite_short = lapply(perf$suite, shorten_suite)
  perf$benchmark = paste0("[",perf$suite_short,"] ", perf$benchmark)
  perf <- normalize(perf, T, "Ř")
  perf_ <- perf[perf$benchmark %in% selection & perf$warm == 1, ]

  ss <<- perf_  # TODO debugging, remove
  
  perf_plot <-
    ggplot(perf_, aes(benchmark, speedup, fill=experiment)) +
    geom_boxplot() +
    geom_hline(yintercept=1) +
    labs(y = "Speedup") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  save_graph(perf_plot, "perf")
}
H3.2a()
```



```{r echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
H3.2b <- function() {
  perf <- fetch_jobs(performanceExperiment)
  perf$experiment = factor(perf$experiment, c("Ř", "Ř-strict"))
  perf$benchmark_ = perf$benchmark
  perf$suite_short = lapply(perf$suite, shorten_suite)
  perf$benchmark = paste0("[",perf$suite_short,"] ", perf$benchmark)
  perf <- normalize(perf, T, "Ř")
  perf_ <- perf[perf$benchmark %in% selection & perf$warm == 1, ]

  ss <<- perf_  # TODO debugging, remove
  
  perf_plot <-
    ggplot(perf_, aes(benchmark, speedup, fill=experiment)) +
    geom_boxplot() +
    geom_hline(yintercept=1) +
    labs(y = "Speedup") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  save_graph(perf_plot, "perf")
}
H3.2b()
```



```{r echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
H3.3 <- function() {
  perf <- fetch_jobs(performanceExperiment)
  perf$experiment = factor(perf$experiment, c("Ř", "Ř-strict"))
  perf$benchmark_ = perf$benchmark
  perf$suite_short = lapply(perf$suite, shorten_suite)
  perf$benchmark = paste0("[",perf$suite_short,"] ", perf$benchmark)
  perf <- normalize(perf, T, "Ř")
  perf_ <- perf[perf$benchmark %in% selection & perf$warm == 1, ]

  ss <<- perf_  # TODO debugging, remove
  
  perf_plot <-
    ggplot(perf_, aes(benchmark, speedup, fill=experiment)) +
    geom_boxplot() +
    geom_hline(yintercept=1) +
    labs(y = "Speedup") +
    theme(axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1))

  save_graph(perf_plot, "perf")
}
H3.3()
```




```{r echo=FALSE, warning=FALSE, message=FALSE, eval=FALSE}
plot_bm <- function(title, d, warmup, trend=FALSE, log=TRUE, relative=FALSE, ncol=6, hide_legend=F, baseline=relative, scales="free") {
  multipleBm <- length(unique(d$benchmark)) > 1
  sz <- if (multipleBm) 0.8 else 1.5
  
  nwarm = WARM_RUNS
  ncold = COLD_RUNS
  cold_position <- position_var_nudge(l_off = nwarm)
  warm_position <- if (warmup)
    position_var_nudge(r_off = ncold)
  else
    position_var_nudge()
  
  experiments = length(unique(d$experiment))
  
  
  print(ggplot(d, if (relative) aes(experiment, speedup) else aes(experiment, ms)) +
          (if (multipleBm) facet_wrap(~benchmark, ncol=ncol, strip.position="left", scales=scales)) +
          # (if (!multipleBm) theme_minimal()) +
          
          #  (if (experiments == 2)
          #    scale_color_manual(values = scales::viridis_pal()(10)[c(1,9)])
          #  else
          #    scale_colour_viridis_d()) +
          
          theme(axis.title.x=element_blank(),
                axis.text.x=element_blank(),
                axis.title.y=element_blank(),
                axis.ticks=element_blank(),
                legend.position=(if (hide_legend) "none" else "right"),
                (if (multipleBm) panel.grid.minor=element_blank())) +
          
          ggtitle(title) +
          (if (log) scale_y_continuous(trans = "log10")) +
          (if (baseline) geom_hline(yintercept=1, color="red", linetype="dashed", size=1)) +
          
          (if (warmup)
            geom_point(position=cold_position,
                       shape=4,
                       size=sz,
                       color="black",
                       aes(row=row),
                       data=subset(d, warm == 0))) +
          geom_point(position=warm_position,
                     size=sz,
                     aes(color=experiment, row=row),
                     data=subset(d, warm == 1)) +
          geom_boxplot(outlier.shape = NA, aes(color=experiment), data=subset(d, warm == 1)) +
          
          (if (trend != F && !multipleBm) {
            # TODO find out how to draw trend on multiple ones
            st = trend[trend$benchmark == d$benchmark_, ]$prediction_start[1]
            ed = trend[trend$benchmark == d$benchmark_, ]$prediction_end[1]
            printResult(paste0("TrendLineEnds", d$benchmark_[1]), (ed-1)*100, 2)
            geom_segment(aes(x = 1, y = st, xend = 7, yend = ed), size=0.01, color="blue")
          })
        # theme(axis.text.x = element_text(angle=90))
  )
}

plot_all_bms <- function(name, data, warmup, trend=FALSE, relative=F) {
  dir.create(name, showWarnings=F)
  setwd(name)
  for (b in unique(data$benchmark)) {
    d <- data[data$benchmark == b,]
    bm_file_name <- paste0(d$suite[1], "-", b, ".pdf")
    bm_title <- paste0("[", d$suite[1], "] ", b)
    cairo_pdf(bm_file_name)
    plot_bm(bm_title, d, warmup=warmup, trend=trend, log=T, relative=relative)
    dev.off()
  }
  setwd("..")
}

plot_by_suite <- function(name, data, warmup, trend=FALSE, removeNormalized=F) {
  dir.create(name, showWarnings=F)
  setwd(name)
  data <- normalize(data, removeNormalized)
  for (s in unique(data$suite)) {
    d <- data[data$suite == s,]
    bm_file_name <- paste0(s, ".pdf")
    cairo_pdf(bm_file_name, width=12, height=12)
    bm_title <- paste0("[", s, "]")
    plot_bm(bm_title, d, warmup, trend, log=T, relative=T)
    dev.off()
  }
  setwd("..")
}

s <- fetch_jobs(specializationExperiment, "Ř")
cmp <- fetch_jobs(vmCompare)
cmp$experiment = factor(cmp$experiment, c("Ř", "FastR", "GNU R"))

shorten_suite = function(s) switch(s, "are-we-fast-r"="awf", "real_thing"="re", "shootout"="sht", "simple"="μ")
pl <- function(models) {
  dir.create("final", showWarnings=F)
  setwd("final")

  cmp$benchmark_ = cmp$benchmark
  s$benchmark_ = s$benchmark
  s$suite_short = lapply(s$suite, shorten_suite)
  s$benchmark = paste0("[",s$suite_short,"] ",s$benchmark)
  cmp$suite_short = lapply(cmp$suite, shorten_suite)
  cmp$benchmark = paste0("[",cmp$suite_short,"] ",cmp$benchmark)

  ncol=4

  s <- normalize(s, F, "level 0")
  cmp <- normalize(cmp, T, "GNU R")

  for (b in c("[sht] spectralnorm")) {
    d <- cmp[cmp$benchmark == b,]
    bm_title <- d$benchmark
    cairo_pdf(paste0(d$benchmark_, ".pdf"), width=4, height=4)
    plot_bm(bm_title, d, warmup=T, trend=F, log=T, relative=T)
    dev.off()

    d <- s[s$benchmark == b,]
    bm_title <- d$benchmark
    cairo_pdf(paste0(d$benchmark_, "-specialization.pdf"), width=4, height=4)
    plot_bm(bm_title, d, warmup=F, trend=models, log=F, relative=T)
    dev.off()
  }

  s_ = s[s$benchmark %in% selection, ]
  cmp_ = cmp[cmp$benchmark %in% selection, ]

  bms = length(unique(s$benchmark))
  bms_ = length(unique(s_$benchmark))

  cairo_pdf("performance.pdf", width=ncol*2, height=2*bms_/ncol)
  plot_bm(NULL, cmp_, warmup=F, relative=T, ncol=ncol, hide_legend=T)
  dev.off()

  cairo_pdf("specialization.pdf", width=ncol*2.5, height=2.5*bms_/ncol)
  plot_bm(NULL, s_, warmup=F, trend=F, log=F, relative=T, ncol=ncol, hide_legend=T, baseline=F)
  dev.off()

  ncol=3
  nrow=4
  pages = (bms / (ncol*nrow))

  for (page in 0:pages) {
    pos = ncol*nrow*page + 1
    npos = ncol*nrow*(page+1)
    sel = unique(cmp$benchmark)[pos:npos]
    d = cmp[cmp$benchmark %in% sel, ]
    bms = length(unique(d$benchmark))
    filename = paste0("performance-appendix-",page,".pdf")
    cairo_pdf(filename, width=ncol*2, height=2.1*bms/ncol)
    plot_bm(NULL, d,
            warmup=T, relative=T, ncol=ncol, hide_legend=T)
    dev.off()
  }

  for (page in 0:pages) {
    pos = ncol*nrow*page + 1
    npos = ncol*nrow*(page+1)
    sel = unique(s$benchmark)[pos:npos]
    d = s[s$benchmark %in% sel, ]
    bms = length(unique(d$benchmark))
    filename = paste0("specialization-appendix-",page,".pdf")
    cairo_pdf(filename, width=ncol*2, height=2.1*bms/ncol)
    plot_bm(NULL, d,
            warmup=F, trend=F, relative=T, log=F, ncol=ncol, hide_legend=T, baseline=F)
    dev.off()
  }

  setwd("..")
}

sumaryCmp <- function(cmp) {
  sumary <- data.frame()
  sumary$suite = character()
  sumary$benchmark = character()
  sumary$vm = character()
  sumary$time = numeric()

  d1 = cmp
  for (su in unique(d1$suite)) {
    d2 <- d1[d1$suite == su, ]
    for (b in unique(d2$benchmark)) {
      d3 <- d2[d2$benchmark == b, ]
      for (e in unique(d3$experiment)) {
        m = median(d3[d3$experiment == e, ]$ms)
        sumary[nrow(sumary)+1,] = list(
          suite=su, benchmark=b, vm=e, time=m)
      }
    }
  }
  sumary$speedup_gnur = NA
  sumary$speedup_fastr = NA

  for (b in unique(cmp$benchmark)) {
    sumary[sumary$vm == "Ř" & sumary$benchmark == b,]$speedup_gnur =
      sumary[sumary$vm == "GNU R" & sumary$benchmark == b,]$time /
      sumary[sumary$vm == "Ř" & sumary$benchmark == b,]$time
    if (nrow(sumary[sumary$vm == "FastR" & sumary$benchmark == b,]))
      sumary[sumary$vm == "Ř" & sumary$benchmark == b,]$speedup_fastr =
        sumary[sumary$vm == "FastR" & sumary$benchmark == b,]$time /
        sumary[sumary$vm == "Ř" & sumary$benchmark == b,]$time
  }
  sumary = sumary[sumary$vm == "Ř",]

  cat("% Performance comparison GNU R\n")
  for (su in unique(cmp$suite)) {
    mi_g = min(sumary[sumary$suite == su,]$speedup_gnur)
    ma_g = max(sumary[sumary$suite == su,]$speedup_gnur)
    me_g = geometric.mean(sumary[sumary$suite == su,]$speedup_gnur)
    mi_f = min(sumary[sumary$suite == su,]$speedup_fastr, na.rm=T)
    ma_f = max(sumary[sumary$suite == su,]$speedup_fastr, na.rm=T)
    me_f = geometric.mean(sumary[sumary$suite == su,]$speedup_fastr, na.rm=T)
    printResult(paste0(shorten_suite(su),"GnurMin"), mi_g)
    printResult(paste0(shorten_suite(su),"GnurMax"), ma_g)
    printResult(paste0(shorten_suite(su),"GnurMed"), me_g)
  }
  mi_g = min(sumary[sumary$suite != "simple",]$speedup_gnur)
  ma_g = max(sumary[sumary$suite != "simple",]$speedup_gnur)
  me_g = geometric.mean(sumary[sumary$suite != "simple",]$speedup_gnur)
  printResult("overallGnurMin", mi_g)
  printResult("overallGnurMax", ma_g)
  printResult("overallGnurMed", me_g)
  printResult("overallGnurMinRounded", mi_g, 1)
  printResult("overallGnurMaxRounded", ma_g, 1)
  printResult("overallGnurMedRounded", me_g, 2)

  cat("% Performance comparison FastR\n")
  for (su in unique(cmp$suite)) {
    mi_g = min(sumary[sumary$suite == su,]$speedup_gnur)
    ma_g = max(sumary[sumary$suite == su,]$speedup_gnur)
    me_g = geometric.mean(sumary[sumary$suite == su,]$speedup_gnur)
    mi_f = min(sumary[sumary$suite == su,]$speedup_fastr, na.rm=T)
    ma_f = max(sumary[sumary$suite == su,]$speedup_fastr, na.rm=T)
    me_f = geometric.mean(sumary[sumary$suite == su,]$speedup_fastr, na.rm=T)
    printResult(paste0(shorten_suite(su),"FastrMin"), mi_f)
    printResult(paste0(shorten_suite(su),"FastrMax"), ma_f)
    printResult(paste0(shorten_suite(su),"FastrMed"), me_f)
  }
  mi_f = min(sumary[sumary$suite != "simple",]$speedup_fastr, na.rm=T)
  ma_f = max(sumary[sumary$suite != "simple",]$speedup_fastr, na.rm=T)
  me_f = geometric.mean(sumary[sumary$suite != "simple",]$speedup_fastr, na.rm=T)
  printResult("overallFastrMin", mi_f)
  printResult("overallFastrMax", ma_f)
  printResult("overallFastrMed", me_f)
  printResult("overallFastrMinRounded", mi_f, 1)
  printResult("overallFastrMaxRounded", ma_f, 1)
  printResult("overallFastrMedRounded", me_f, 2)
}

sumarySpecialization <- function(s, kind) {
  cat(paste0("% contextual dispatch experiment H0 ",kind,"\n"))
  data = normalize(s, removeNormalized=F, "level 0")
  data = data[data$warm == 1,]
  data$speedup = data$speedup - 1.0
  data$experiment = factor(data$experiment,
                           levels=c("level 0", "level 1", "level 2", "level 3", "level 4", "level 5", "level 6"))
  data$experiment = as.integer(data$experiment)-1

  sumary2 <- data.frame()
  sumary2$speedup=numeric()
  sumary2$accept =numeric()
  sumary2$reject =numeric()
  sumary2$avg =numeric()

  models = data.frame(benchmark=unique(data$benchmark))
  models$prediction_start = 0
  models$prediction_end = 0

  speedups <- c(0.95, 0.98, 1.0, 1.02, 1.05, 1.1, 1.2)
  for (speedup_ in speedups)
    sumary2[nrow(sumary2)+1,]=list(
      speedup=speedup_, accept=0, reject=0, avg=0)

  hypothesisTest <- function(simple) {
    d1 <- data
    for (su in unique(d1$suite)) {
      if (su == "simple" && !simple)
        next()
      if (su != "simple" && simple)
        next()
      d2 <- d1[d1$suite == su, ]
      for (b in unique(d2$benchmark)) {
        d3 = d2[d2$benchmark == b, ]
        lm_ <- lm(speedup ~ experiment, data = d3)
        prediction_start = predict(lm_, list(experiment=0), interval="confidence")
        prediction_end = predict(lm_, list(experiment=6), interval="confidence")
        models[models$benchmark == b,]$prediction_start = prediction_start[2] +1
        models[models$benchmark == b,]$prediction_end = prediction_end[2] +1
        p = summary(lm_)$coefficients[,4]
        for (speedup_ in speedups) {
          prediction = 1+predict(lm_, list(experiment=6), interval="confidence")
          h = prediction[2] > speedup_
          avg = prediction[1] > speedup_
          if (speedup_ == 0.95 && !h)
            cat(paste0("% Bad slowdown in :", b, "\n"))
          if (speedup_ == 1.1 && h)
            cat(paste0("% Large speedup in :", b, "\n"))
          if (avg) {
            sumary2[sumary2$speedup == speedup_,]$avg =
              sumary2[sumary2$speedup == speedup_,]$avg+1
          }
          if (h) {
            sumary2[sumary2$speedup == speedup_,]$reject =
              sumary2[sumary2$speedup == speedup_,]$reject+1
          } else {
            sumary2[sumary2$speedup == speedup_,]$accept =
              sumary2[sumary2$speedup == speedup_,]$accept + 1
          }
        }
      }
    }
    for (speedup_ in speedups) {
      r = sumary2[sumary2$speedup == speedup_,]
      s = switch(as.character(speedup_),
                 "0.95"="MinusFive", "0.98"="MinusTwo", "1"="One", "1.02"="PlusTwo", "1.05"="PlusFive", "1.1"="PlusTen", "1.2"="PlusTwenty")
      printResult(paste0("HReject", s, (if (simple) "Mu" else ""), kind), r$reject)
    }
    invisible(models)
  }
  cat("% for mu\n")
  hypothesisTest(T)
  cat("% for others\n")
  hypothesisTest(F)
}

printResult("BenchmarksInTotal",      length(unique(s$benchmark)), space=T)
printResult("BenchmarksInTotalMu",    length(unique(s[s$suite == "simple",]$benchmark)), space=T)
printResult("BenchmarksInTotalNonMu", length(unique(s[s$suite != "simple",]$benchmark)), space=T)
sumaryCmp(cmp)
m = sumarySpecialization(s, "")
cat("% some key numbers from the plots\n")
pl(m)
```
